Improving the communication and transparency of stock assessment using interactive

### 2018-11-09 manuscript.Rmd
Rough draft of data section intro; set-up for tagging dashboard example

The field of fisheries research has a long history of collecting and synthesizing a plethora of data using state-of-the-art computational technologies [@megrey2008]. Fisheries science has evolved alongside the rise of the computing era, and as tools improve so does the management advice provided by fisheries scientists. This management advice generally culminates from a stock assessment process whereby various sources of data and statistical methods are used to model the history, current status and future status of one or more fish stocks [@hilborn1992]. Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated assessment models [@maunder2013]. This means that analyses that were once carried out independently on different data sets can now be integrated into the one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015]. From a computational perspective, analyses of a variety of large data sets has never been easier. However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers. Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever expanding array of data inputs and model outputs. First, we focus on data sets that are commonly used in stock assessment and provide one example where an interactive maps were used to simplify the detailed exploration of data from a long-term tagging program. Second, we focus on the modeling aspect of stock assessment and, again, we use an example to demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model.

## Data

The data sets used in stock assessments are constantly growing. This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs. As such, stock assessment biologists often have to manage large volumes of data from a variety of sources. For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments. These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and to synthesize the results.

## Models

@@

>>>>

The field of fisheries research has a long history of collecting and synthesizing a plethora of data using state-of-the-art computational technologies [@megrey2008]. Fisheries science has evolved alongside the rise of the computing era, and as tools improve so does the management advice provided by fisheries scientists. This management advice generally culminates from a stock assessment process whereby various sources of data and statistical methods are used to model the history, current status and future status of one or more fish stocks [@hilborn1992]. Recent advances in computational technologies have streamlined the process of stock assessment by facilitating the development and application of integrated assessment models [@maunder2013]. This means that analyses that were once carried out independently on different data sets can now be integrated into one analysis by using statistical modeling tools such as JAGS [@plummer2003], AD Model Builder [@fournier2012] and Template Model Builder [@kristensen2015]. From a computational perspective, analyses of a variety of large data sets has never been easier. However, from a human perspective, contemporary stock assessment biologists are faced with the challenge of managing data from a variety of sources and also understanding the algorithms that convert these data to advice for fisheries managers. Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever expanding array of data inputs and model outputs. First, we focus on data that are commonly used in stock assessment and provide one example where interactive maps were used to simplify the detailed exploration of data from a long-term tagging program. Second, we focus on the modeling aspect of stock assessment and, again, we use an example to demonstrate how dynamic and interactive data visualizations can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to the two general steps in the stock assessment process: 1) knowing the data, and 2) data modeling.

## Knowing the data

The data sets used in stock assessments are constantly growing. This growth in data either stems from the continuation of long-term monitoring efforts or from the addition of new monitoring programs. As such, stock assessment biologists often have to manage large volumes of data from a variety of sources. For instance, time series of reported landings and catch-at-age are "fishery-dependent" data that are frequently used in stock assessments. These data are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are multidimensional and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow the construction of visual representations of the data that demonstrates large-scale patterns while also allowing the detailed exploration of the underlying data. For instance, the ability to zoom in on features or areas of interest and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of the data. This is exemplified by a recently developed interactive mapping tool developed for the exploration of a long-term tagging study.

- *Greg, can you introduce this tool, providing some info on the tagging database and the tools used to make the tool (Shiny, leaflet)?*
- *TODO: the intro paragraph to this section could use some references if it is necessary*

## Data modeling

@@

- *TODO: tweak this intro paragraph and work up sugue to NCAM and the dashboard*
